---
title: "Covid91 vaccine study Final2024"
author: "Sarah Castellon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



## Males

Question: How effective is the vaccine on males?

##Method 

$H_0$ the vaccine is not effective in preventing covid 91 in males  
$H_A$ the vaccine is effective in preventing covid 91 in males  

Infected~Sex  
Categorical ~ Categorical   
We will use a helper file Cat~Cat  

```{r}
Male <- FinalDat[FinalDat$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```
## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

## Inferential  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

## Females

Question:How effective is the vaccine on females?

##Method 

$H_0$ the vaccine is not effective in preventing covid 91 in females  
$H_A$ the vaccine is effective in preventing covid 91 in females  

Infected~Sex  
Categorical ~ Categorical   
We will use a helper file Cat~Cat    

```{r}
Female <- FinalDat[FinalDat$sex=="F",]
```

```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```
The bar chart shows that you are less likely to contract covid 91 if your received the vaccine.   

## Numerical Results

```{r}
table2 <- xtabs(~infected + treatment, data=Female)
rowPerc(table2)
colPerc(table2)
```

```{r}
chisq.test(table2)
chisqtestGC(table2)
```

## LGBTQ

Question:How effective is the vaccine on LGBTQ?

##Method 

$H_0$ the vaccine is not effective in preventing covid 91 in LGBTQ  
$H_A$ the vaccine is effective in preventing covid 91 in LGBTQ  

Infected~LGBTQ  
Categorical ~ Categorical   
We will use a helper file Cat~Cat    


## Druggies

Question:How effective is the vaccine on Druggies?

##Method 

$H_0$ the vaccine is not effective in preventing covid 91 in Druggies  
$H_0A$ the vaccine is effective in preventing covid 91 in Druggies  

Infected~DrugUser  
Categorical ~ Categorical   
We will use a helper file Cat~Cat   

# Overall Results and Conclusions