---
title: "Covid91 vaccine study Final2024"
author: "Sarah Castellon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


# Males

Question: How effective is the vaccine on males?

## Method 

$H_0$ the vaccine is not effective in preventing covid 91 in males  
$H_A$ the vaccine is effective in preventing covid 91 in males  

Infected~Sex  
Categorical ~ Categorical   
We will use a helper file Cat~Cat  

## Graphical  

```{r}
Male <- FinalDat[FinalDat$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```
The bar chart shows that males who recieved the vaccine are less likely to contact covid91.  

## Numerical 

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```
The numerical data shows that within the placebo group 73.80% of males got infected with covid 91, while of the males who received the vaccine only 26.20% contracted covid91. Therefore this indicates that the males who received the vaccine significantly reduced the risk of contracting covid91.


## Inferential  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Square p value is .00000000000000022 Fisher exact test shows the p-value is .00000000000000022 which is lower than 0.05 so we reject the null hypothesis.  

## Conclusion 

Regarding the males the bar chart shows you are less likely to contract covid91 if you received the vaccine. The numerical data shows that within the placebo group 73.80% of males got infected with covid 91, while the males who received the vaccine only 26.20% contracted covid 91. Therefore, this indicates that the males who received the vaccine significantly reduced the risk of contracting covid 91.The Fisher exact test shows the p-value is .00000000000000022 which is lower than 0.05 so we reject the null hypothesis.    

# Females

Question:How effective is the vaccine on females?

## Method 

$H_0$ the vaccine is not effective in preventing covid 91 in females  
$H_A$ the vaccine is effective in preventing covid 91 in females  

Infected~Sex  
Categorical ~ Categorical   
We will use a helper file Cat~Cat    

## Graphical  

```{r}
Female <- FinalDat[FinalDat$sex=="F",]
```

```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```
The bar chart shows that females that received the vaccine are less likely to contract covid 91.  

## Numerical 

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 64.93% of females got infected with covid 91, while of the females who received the vaccine only 35.07% contracted covid91. Therefore this indicates that the females who received the vaccine significantly reduced the risk of contracting covid91.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Square test show the p value is .00000001914 and the  Fisher exact test shows the p-value is .00000001914 which is lower than 0.05 so we reject the null hypothesis.


## Conclusion

The bar chart shows that females that received the vaccine are less likely to contract covid 91. The numerical data shows that within the placebo group 64.93% of females got infected with covid 91, while of the females who received the vaccine only 35.07% contracted covid 91. Therefore this indicates that the females who received the vaccine significantly reduced the risk of contracting covid 91. The Fisher exact test shows the p-value is .00000001914 which is lower than 0.05 so we reject the null hypothesis.

# LGBTQ

Question:How effective is the vaccine on LGBTQ?

## Method 

$H_0$ the vaccine is not effective in preventing covid 91 in LGBTQ  
$H_A$ the vaccine is effective in preventing covid 91 in LGBTQ  

Infected~LGBTQ(gay)  
Categorical ~ Categorical   
We will use a helper file Cat~Cat    

## Graphical

```{r}
LBGTQ <- FinalDat[FinalDat$LGBTQ=='gay',]
```

```{r}
barchartGC(~infected + treatment,data=LBGTQ)
barchartGC(~infected + treatment,data=LBGTQ, type="percent")

```
The bar chart shows that members of the LBGTQ community are less likely to contract covid 91 if your received the vaccine.   

## Numberical

```{r}
table1 <- xtabs(~infected + treatment, data=LBGTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 80% of the LBGTQ community members got infected with covid 91, while of the LBGTQ community members who received the vaccine only 20% contracted covid 91.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi square test p value is .000002266 and the Fisher exact test shows the p-value is .0000007269 which is lower than 0.05 so we reject the null hypothesis.

## Conclusion  

The bar chart shows that members if the LBGTQ community are less likely to contract covid 91 if your received the vaccine. The numerical data shows that within the placebo group 80% of the LBGTQ community members got infected with covid 91, while of the LBGTQ community members who received the vaccine only 20% contracted covid 91.The Fisher exact test shows the p-value is .0000007269 which is lower than 0.05 so we reject the null hypothesis.  

# Druggies

Question:How effective is the vaccine on Druggies?

## Method 

$H_0$ the vaccine is not effective in preventing covid 91 in Druggies  
$H_0A$ the vaccine is effective in preventing covid 91 in Druggies  

Infected~DrugUser  
Categorical ~ Categorical   
We will use a helper file Cat~Cat   

## Graphical

```{r}
DrugUser <- FinalDat[FinalDat$DrugUser=='yes',]
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

The bar chart shows that members of the DrugUsers community are more likely to contract covid 91 if your received the vaccine. 

## Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 8.33% of the DrugUser community members got infected with covid 91, while of the DrugUser community members who received the vaccine only 91.76% contracted covid 91. This indicates those who received that vaccine had a significant increase of contracting covid 91.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi square test p value is .0000000001414 and and the Fisher exact test shows the p-value is .000000000003261 which is lower than 0.05 so we reject the null hypothesis.

## Conclusion  

The bar chart shows that members of the DrugUsers community are more likely to contract covid 91 if your received the vaccine. The numerical data shows that within the placebo group 8.33% of the DrugUser community members got infected with covid 91, while of the DrugUser community members who received the vaccine only 91.76% contracted covid 91. This indicates those who received that vaccine had a significant increase of contracting covid 91.The Chi square test p value is .0000000001414 and and the Fisher exact test shows the p-value is .000000000003261 which is lower than 0.05 so we reject the null hypothesis. This data shows that DrugUsers are far more likely to contract covid 91 if they received th vaccine. 


# Overall Results and Conclusions

Overall the data reflects that receiving the vaccine generally help with preventing the contraction of covid 91 if you are male, female or a members of the LBGTQ community. However, in the DrugUser  community those who received the vaccine had a higher risk of contracting covid 91. 